Spatial Coordination Of Stomatal Patterning Between Leaf Surfaces In Amphistomatous \emph{Arabidopsis thaliana} Incurs No Photosynthetic Advantage

### 2023-09-28 ms/ms.Rmd
edited single surface analysis methods

*Arabidopsis thaliana* plants were grown in three different light environments: low light (50 PAR), medium light (100 PAR), and high light (200 PAR) following the light standards from (CITE). Once leaves were mature, we captured images of the abaxial and adaxial leaf surfaces using xxx bifocal microscope. This microscope allows the capture of two focal lengths that are spatially correlated such that pixel (1,1) of the abaxial surface image is directly above pixel (1,1) of the adaxial surface image, and so on... Images were 512 by 512 pixels and covered X leaf surface area. We captured 132 images in total, making 66 abaxial-adaxial image pairs. The position of all stomata were recorded using ImageJ [@schneider_nih_2012].

## Single Surface Analysis

@@

\begin{equation}\label{eq:emd}
  \mathrm{\overline{D}_E} = \frac{0.5}{\sqrt{A_l/n_\text{stomata}}}
\end{equation}

$A_l$ is leaf area and $n_\text{stomata}$ the number of stomata. $\overline{D}_E$ is the theoretical average distance to the nearest neighbor of each stomate if stomata were randomly distributed (CITE; Clark and Evans, 1954). And $\overline{D}_O$ calculated for each synthetic data set is:

\begin{equation}\label{eq:omd}
  \mathrm{\overline{D}_O} = \frac{\sum_{i=1}^{n_\text{stomata}}d_i}{n_\text{stomata}}
\end{equation}

$d_i$ is the distance between $\text{stomate}_i$ and its nearest neighbor. We calculated $\mathrm{NNI}$ using the R package **spatialEco** version `r packageVersion("spatialEco")`.

@@

We calculated the dispersion index $\mathrm{DI}$) which varies from zero to one. Zero is uniformly random and one is ideally dispersed:

@@

Uniformly distributed stomata on a leaf with stomatal density ($D_S$) maximize $\overline{D}_O$ in an equilateral triangle pattern with side lengths ($s$) equal to two times the incircle radius or apothem, $a$, of a regular hexagon with area ($A_\text{hex}$) equal to the area of the leaf ($A_l$) divided by the number of stomata ($n_\text{stomata}$) on the leaf according to Eq. \ref{eq:eq2}:

\begin{equation}\label{eq:eq2}
  s = \frac{\sqrt{2}} {3^{1/4} \sqrt{n_\text{stomata} / A_l}}
\end{equation}

As all sides of an equilateral triangle have the same length, the average nearest neighbor distance for a given $D_S$ is maximized and each stomate occupies the same hexagonal area ($A_\text{hex}$). The closer to one $DI$ is, the more uniform an area each of its stomata supply with CO$_2$ during photosynthesis. We tested whether light treatment affects $\mathrm{DI}$ and $D_S$ using analysis of variance (ANOVA).

Finally, to test our hypothesis that stomatal length is modulated by the area of the leaf to which it supplies CO$_2$, we examined the relationship between stomatal zone area and stomatal length using a Bayesian generalized non-linear multilevel model with the R package **brms** version `r packageVersion("brms")`. Stomatal zone area was calculated using voronoi tessellation (e.g. Fig. \ref{fig:example}). The stomatal zone area, $S_\text{area}$, is the region of the leaf surface whose distance to stomate, $S$, is less than the distance to any other stomate, $S$. Stomatal length was measured in `ImageJ` [@schneider_nih_2012].

>>>>

*Arabidopsis thaliana* plants were grown in three different light environments: low light (50 PAR), medium light (100 PAR), and high light (200 PAR)... Once leaves were mature, we captured images of the abaxial and adaxial leaf surfaces using xxx bifocal microscope. This microscope allows the capture of two focal lengths that are spatially correlated such that pixel (1,1) of the abaxial surface image is directly above pixel (1,1) of the adaxial surface image, and so on... Images were 512 by 512 pixels and covered X leaf surface area. We captured 132 images in total, making 66 abaxial-adaxial image pairs. The position of all stomata were recorded using ImageJ [@schneider_nih_2012].

## Single surface analyses

@@

\begin{equation}\label{eq:emd}
  \overline{D}_E = \frac{0.5}{\sqrt{A_\text{leaf} / n_\text{stomata}}}
\end{equation}

\noindent $A_\text{leaf}$ is leaf area visible in the sampled field and $n_\text{stomata}$ the number of stomata. $\overline{D}_E$ is the theoretical average distance to the nearest neighbor of each stomate if stomata were uniformly randomly distributed [@clark_distance_1954]. And $\overline{D}_O$ calculated for each synthetic data set is:

\begin{equation}\label{eq:omd}
  \overline{D}_O = \frac{\sum_{i=1}^{n_\text{stomata}}d_i}{n_\text{stomata}}
\end{equation}

$d_i$ is the distance between $\text{stomate}_i$ and its nearest neighbor. We calculated $\mathrm{NNI}$ using the *R* package **spatialEco** version `r packageVersion("spatialEco")` [@evans_spatialeco_2023].

@@

We developed a dispersion index $\mathrm{DI}$ to quantify how close observed stomatal distributions are to random uniform versus maximally dispered in an equilateral triangular grid. $\mathrm{DI}$ varies from zero to one, where zero is uniformly random and one is ideally dispersed:

@@

\noindent $\mathrm{NNI}$ is calculated for each sample image as described above; $\text{median}(\mathrm{NNI_{random}})$ and $\text{median}(\mathrm{NNI_{uniform}})$ are calculated from the dry og synthetic data specific to each sample image as described above. We tested whether light treatment affects $\mathrm{DI}$ and $D_S$ using analysis of variance (ANOVA).

<!-- CDM: I am not sure this part is needed
Uniformly distributed stomata on a leaf with stomatal density ($D_S$) maximize $\overline{D}_O$ in an equilateral triangle pattern with side lengths ($s$) equal to two times the incircle radius or apothem, $a$, of a regular hexagon with area ($A_\text{hex}$) equal to $A_\text{leaf}$ divided by the number of stomata ($n_\text{stomata}$) on the leaf according to Eq. \ref{eq:eq2}:

\begin{equation}\label{eq:eq2}
  s = \frac{\sqrt{2}} {3^{1/4} \sqrt{n_\text{stomata} / A_\text{leaf}}}
\end{equation}

As all sides of an equilateral triangle have the same length, the average nearest neighbor distance for a given $D_S$ is maximized and each stomate occupies the same hexagonal area ($A_\text{hex}$). As $DI$ approaches 1, the more uniform an area each of its stomata supply with CO$_2$ during photosynthesis. 
 -->

Finally, we examined the relationship between stomatal zone area and stomatal length using a Bayesian generalized non-linear multilevel model with the R package **brms** version `r packageVersion("brms")` [@burkner_brms_2017; @burkner_advanced_2018]. Stomatal zone area was calculated using Voronoi tessellation (e.g. Fig. \ref{fig:example}). The stomatal zone area, $S_\text{area}$, is the region of the leaf surface whose distance to stomate, $S$, is less than the distance to any other stomate, $S$. Stomatal length was measured in `ImageJ` [@schneider_nih_2012].

### 2023-09-28 ms/ms.tex
edited single surface analysis methods

To maximize CO\(_2\) supply from the stomatal pore to chloroplasts,
stomata should be uniformly distributed in an equilateral triangular
grid on the leaf surface so as to minimize stomatal number and CO\(_2\)
diffusion path length
(\protect\hyperlink{ref-parkhurst_diffusion_1994}{Parkhurst 1994}). As
the diffusion rate of CO\(_2\) though liquid is approximately
\(10^4\times\) slower than CO\(_2\) diffusion through air, mesophyll
resistance is generally thought to be primarily limited by liquid
diffusion (\protect\hyperlink{ref-aalto_three-dimensional_2002}{Aalto
and Juurola 2002}; \protect\hyperlink{ref-evans_resistances_2009}{Evans
et al. 2009}), but diffusion through the IAS has also been shown to be a
rate limiting process because the tortuous, disjunct nature of the IAS
can greatly increase diffusion path lengths
(\protect\hyperlink{ref-harwood_understanding_2021}{Harwood,
Théroux‐Rancourt, and Barbour 2021}). Additionally, tortuosity is higher
in horizontal directions (parallel to leaf surface) than vertical
directions (perpendicular to leaf surface) because of the cylindrical
shape and vertical arrangement of pallisade mesophyll cells
(\protect\hyperlink{ref-earles_beyond_2018}{Earles et al. 2018};
\protect\hyperlink{ref-harwood_understanding_2021}{Harwood,
Théroux‐Rancourt, and Barbour 2021}). However, the ratio of lateral to
vertical diffusion rate is still largely unknown
(\protect\hyperlink{ref-morison_lateral_2005}{Morison et al. 2005};
\protect\hyperlink{ref-pieruschka_lateral_2005}{R. Pieruschka 2005};
\protect\hyperlink{ref-pieruschka_lateral_2006}{Roland Pieruschka et al.
2006}). Depending on the thickness of the leaf, porosity of the leaf
mesophyll, tortuosity of the IAS, and lateral to vertical diffusion rate
ratio, minimizing diffusion path length for CO\(_2\) via optimally
distributed stomata may yield significant increases in CO\(_2\) supply
for photosynthesis and higher \(A_\text{max}\).

@@

\emph{Arabidopsis thaliana} plants were grown in three different light
environments: low light (50 PAR), medium light (100 PAR), and high light
(200 PAR) following the light standards from (CITE). Once leaves were
mature, we captured images of the abaxial and adaxial leaf surfaces
using xxx bifocal microscope. This microscope allows the capture of two
focal lengths that are spatially correlated such that pixel (1,1) of the
abaxial surface image is directly above pixel (1,1) of the adaxial
surface image, and so on\ldots{} Images were 512 by 512 pixels and
covered X leaf surface area. We captured 132 images in total, making 66
abaxial-adaxial image pairs. The position of all stomata were recorded
using ImageJ (\protect\hyperlink{ref-schneider_nih_2012}{Schneider,
Rasband, and Eliceiri 2012}).

\hypertarget{single-surface-analysis}{%
\subsection{Single Surface Analysis}\label{single-surface-analysis}}

@@

\begin{equation}\label{eq:emd}
  \mathrm{\overline{D}_E} = \frac{0.5}{\sqrt{A_l/n_\text{stomata}}}
\end{equation}

\(A_l\) is leaf area and \(n_\text{stomata}\) the number of stomata.
\(\overline{D}_E\) is the theoretical average distance to the nearest
neighbor of each stomate if stomata were randomly distributed (CITE;
Clark and Evans, 1954). And \(\overline{D}_O\) calculated for each
synthetic data set is:

\begin{equation}\label{eq:omd}
  \mathrm{\overline{D}_O} = \frac{\sum_{i=1}^{n_\text{stomata}}d_i}{n_\text{stomata}}
\end{equation}

\(d_i\) is the distance between \(\text{stomate}_i\) and its nearest
neighbor. We calculated \(\mathrm{NNI}\) using the R package
\textbf{spatialEco} version 2.0.1.

@@

We calculated the dispersion index \(\mathrm{DI}\)) which varies from
zero to one. Zero is uniformly random and one is ideally dispersed:

@@

Uniformly distributed stomata on a leaf with stomatal density (\(D_S\))
maximize \(\overline{D}_O\) in an equilateral triangle pattern with side
lengths (\(s\)) equal to two times the incircle radius or apothem,
\(a\), of a regular hexagon with area (\(A_\text{hex}\)) equal to the
area of the leaf (\(A_l\)) divided by the number of stomata
(\(n_\text{stomata}\)) on the leaf according to Eq. \ref{eq:eq2}:

\begin{equation}\label{eq:eq2}
  s = \frac{\sqrt{2}} {3^{1/4} \sqrt{n_\text{stomata} / A_l}}
\end{equation}

As all sides of an equilateral triangle have the same length, the
average nearest neighbor distance for a given \(D_S\) is maximized and
each stomate occupies the same hexagonal area (\(A_\text{hex}\)). The
closer to one \(DI\) is, the more uniform an area each of its stomata
supply with CO\(_2\) during photosynthesis. We tested whether light
treatment affects \(\mathrm{DI}\) and \(D_S\) using analysis of variance
(ANOVA).

Finally, to test our hypothesis that stomatal length is modulated by the
area of the leaf to which it supplies CO\(_2\), we examined the
relationship between stomatal zone area and stomatal length using a
Bayesian generalized non-linear multilevel model with the R package
\textbf{brms} version 2.20.1. Stomatal zone area was calculated using
voronoi tessellation (e.g.~Fig. \ref{fig:example}). The stomatal zone
area, \(S_\text{area}\), is the region of the leaf surface whose
distance to stomate, \(S\), is less than the distance to any other
stomate, \(S\). Stomatal length was measured in \texttt{ImageJ}
(\protect\hyperlink{ref-schneider_nih_2012}{Schneider, Rasband, and
Eliceiri 2012}).

@@

\hypertarget{single-surface-analysis-1}{%
\subsection{Single surface analysis}\label{single-surface-analysis-1}}

@@

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-gay_influence_1975}{}}%
Gay, A. P., and R. G. Hurd. 1975. {``{THE} {INFLUENCE} {OF} {LIGHT} {ON}
{STOMATAL} {DENSITY} {IN} {THE} {TOMATO}.''} \emph{New Phytologist} 75
(1): 37--46. \url{https://doi.org/10.1111/j.1469-8137.1975.tb01368.x}.

>>>>

To maximize CO\(_2\) supply from the stomatal pore to chloroplasts,
stomata should be uniformly distributed in an equilateral triangular
grid on the leaf surface so as to minimize stomatal number and CO\(_2\)
diffusion path length
(\protect\hyperlink{ref-parkhurst_diffusion_1994}{Parkhurst 1994}). As
the diffusion rate of CO\(_2\) though liquid is approximately
\(10^4\times\) slower than CO\(_2\) diffusion through air, mesophyll
resistance is generally thought to be primarily limited by liquid
diffusion (\protect\hyperlink{ref-aalto_three-dimensional_2002}{Aalto
and Juurola 2002}; \protect\hyperlink{ref-evans_resistances_2009}{J. R.
Evans et al. 2009}), but diffusion through the IAS has also been shown
to be a rate limiting process because the tortuous, disjunct nature of
the IAS can greatly increase diffusion path lengths
(\protect\hyperlink{ref-harwood_understanding_2021}{Harwood,
Théroux‐Rancourt, and Barbour 2021}). Additionally, tortuosity is higher
in horizontal directions (parallel to leaf surface) than vertical
directions (perpendicular to leaf surface) because of the cylindrical
shape and vertical arrangement of pallisade mesophyll cells
(\protect\hyperlink{ref-earles_beyond_2018}{Earles et al. 2018};
\protect\hyperlink{ref-harwood_understanding_2021}{Harwood,
Théroux‐Rancourt, and Barbour 2021}). However, the ratio of lateral to
vertical diffusion rate is still largely unknown
(\protect\hyperlink{ref-morison_lateral_2005}{Morison et al. 2005};
\protect\hyperlink{ref-pieruschka_lateral_2005}{R. Pieruschka 2005};
\protect\hyperlink{ref-pieruschka_lateral_2006}{Roland Pieruschka et al.
2006}). Depending on the thickness of the leaf, porosity of the leaf
mesophyll, tortuosity of the IAS, and lateral to vertical diffusion rate
ratio, minimizing diffusion path length for CO\(_2\) via optimally
distributed stomata may yield significant increases in CO\(_2\) supply
for photosynthesis and higher \(A_\text{max}\).

@@

\emph{Arabidopsis thaliana} plants were grown in three different light
environments: low light (50 PAR), medium light (100 PAR), and high light
(200 PAR)\ldots{} Once leaves were mature, we captured images of the
abaxial and adaxial leaf surfaces using xxx bifocal microscope. This
microscope allows the capture of two focal lengths that are spatially
correlated such that pixel (1,1) of the abaxial surface image is
directly above pixel (1,1) of the adaxial surface image, and so
on\ldots{} Images were 512 by 512 pixels and covered X leaf surface
area. We captured 132 images in total, making 66 abaxial-adaxial image
pairs. The position of all stomata were recorded using ImageJ
(\protect\hyperlink{ref-schneider_nih_2012}{Schneider, Rasband, and
Eliceiri 2012}).

\hypertarget{single-surface-analyses}{%
\subsection{Single surface analyses}\label{single-surface-analyses}}

@@

\begin{equation}\label{eq:emd}
  \overline{D}_E = \frac{0.5}{\sqrt{A_\text{leaf} / n_\text{stomata}}}
\end{equation}

\noindent \(A_\text{leaf}\) is leaf area visible in the sampled field
and \(n_\text{stomata}\) the number of stomata. \(\overline{D}_E\) is
the theoretical average distance to the nearest neighbor of each stomate
if stomata were uniformly randomly distributed
(\protect\hyperlink{ref-clark_distance_1954}{Clark and Evans 1954}). And
\(\overline{D}_O\) calculated for each synthetic data set is:

\begin{equation}\label{eq:omd}
  \overline{D}_O = \frac{\sum_{i=1}^{n_\text{stomata}}d_i}{n_\text{stomata}}
\end{equation}

\(d_i\) is the distance between \(\text{stomate}_i\) and its nearest
neighbor. We calculated \(\mathrm{NNI}\) using the R package
\textbf{spatialEco} version 2.0.1
(\protect\hyperlink{ref-evans_spatialeco_2023}{J. S. Evans and Murphy
2023}).

@@

We developed a dispersion index \(\mathrm{DI}\) to quantify how close
observed stomatal distributions are to random uniform versus maximally
dispered in an equilateral triangular grid. \(\mathrm{DI}\) varies from
zero to one, where zero is uniformly random and one is ideally
dispersed:

@@

\noindent \(\mathrm{NNI}\) is calculated for each sample image as
described above; \(\text{median}(\mathrm{NNI_{random}})\) and
\(\text{median}(\mathrm{NNI_{uniform}})\) are calculated from the dry og
synthetic data specific to each sample image as described above. We
tested whether light treatment affects \(\mathrm{DI}\) and \(D_S\) using
analysis of variance (ANOVA).

Finally, we examined the relationship between stomatal zone area and
stomatal length using a Bayesian generalized non-linear multilevel model
with the R package \textbf{brms} version 2.20.1
(\protect\hyperlink{ref-burkner_brms_2017}{Bürkner 2017},
\protect\hyperlink{ref-burkner_advanced_2018}{2018}). Stomatal zone area
was calculated using Voronoi tessellation (e.g.~Fig. \ref{fig:example}).
The stomatal zone area, \(S_\text{area}\), is the region of the leaf
surface whose distance to stomate, \(S\), is less than the distance to
any other stomate, \(S\). Stomatal length was measured in
\texttt{ImageJ} (\protect\hyperlink{ref-schneider_nih_2012}{Schneider,
Rasband, and Eliceiri 2012}).

@@

\hypertarget{single-surface-analysis}{%
\subsection{Single surface analysis}\label{single-surface-analysis}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-burkner_brms_2017}{}}%
Bürkner, Paul-Christian. 2017. {``\textbf{Brms} : {An} \emph{r}
{Package} for {Bayesian} {Multilevel} {Models} {Using} \emph{Stan}.''}
\emph{Journal of Statistical Software} 80 (1).
\url{https://doi.org/10.18637/jss.v080.i01}.

\leavevmode\vadjust pre{\hypertarget{ref-burkner_advanced_2018}{}}%
---------. 2018. {``Advanced {Bayesian} {Multilevel} {Modeling} with the
{R} {Package} Brms.''} \emph{The R Journal} 10 (1): 395.
\url{https://doi.org/10.32614/RJ-2018-017}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-clark_distance_1954}{}}%
Clark, Philip J., and Francis C. Evans. 1954. {``Distance to Nearest
Neighbor as a Measure of Spatial Relationships in Populations.''}
\emph{Ecology} 35 (4): 445--53. \url{https://doi.org/10.2307/1931034}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-evans_spatialeco_2023}{}}%
Evans, Jeffrey S., and Melanie A. Murphy. 2023. \emph{{spatialEco}}.
\url{https://github.com/jeffreyevans/spatialEco}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-gay_influence_1975}{}}%
Gay, A. P., and R. G. Hurd. 1975. {``The Influence of Light on Stomatal
Density in the Tomato.''} \emph{New Phytologist} 75 (1): 37--46.
\url{https://doi.org/10.1111/j.1469-8137.1975.tb01368.x}.
